%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% abstract-en.tex
%% NOVA thesis document file
%%
%% Abstract in English
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE abstract-en.tex}%

The \gls{IoT} is transforming monitoring and surveillance across sectors such as manufacturing, agriculture, and smart cities. However, building platforms that handle heterogeneous devices at scale, process data in real time, and remain operational during network outages remains challenging.

This thesis presents the design, implementation, and evaluation of an open and agnostic cloud-based \gls{IoT} event management platform for monitoring and surveillance. The platform adopts a microservice architecture with event-driven communication, using \gls{MQTT} for device telemetry ingestion and Apache Kafka as the internal messaging backbone. A schema-agnostic ingestion model automatically discovers and normalizes telemetry fields from any device without prior configuration. The analytics engine supports rule-based threshold alerting and Z-score anomaly detection over the telemetry stream. A web application provides live dashboards with delivery of telemetry and alerts via WebSocket.

The platform is extended with an edge computing node that provides offline resilience through a write-first data persistence pattern. The edge node performs local telemetry ingestion and \texttt{CRITICAL} rule evaluation, buffering all data in SQLite during connectivity interruptions and synchronizing with the cloud upon reconnection. Asymmetric \gls{JWT} signing (RS256) enables fully offline device authentication on the edge.

The platform was evaluated through five experiments. A baseline performance study measured throughput, latency, and resource utilization across the three ingestion protocols (\gls{MQTT}, \gls{HTTP}, and \gls{CoAP}) under load profiles from 100 to 10\,000 messages per second. A scaling study on the \gls{MQTT} path showed that adding bridge replicas and Kafka partitions increased throughput by 174\%. An offline resilience test confirmed that the edge node preserved all buffered data during a five-minute disconnection. An extensibility test added a Python-based vision service through Kafka without modifying any existing service. A device heterogeneity test with a Raspberry Pi sensor confirmed automatic field discovery across different producer types. The results support the design goals of real-time processing, extensibility, and offline resilience, while identifying the scalability boundaries of a single-node deployment.

\keywords{
	Internet of Things \and
	Microservices \and
	Event-Driven Architecture \and
	Edge Computing \and
	Real-Time Monitoring \and
	MQTT \and
	Apache Kafka
}
