%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% acknowledgements.tex
%% NOVA thesis document file
%%
%% Text with acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE acknowledgements.tex}%

\begin{ntacknowledgements}
I would like to express my gratitude to all those who contributed, directly or indirectly, to the completion of this dissertation.

I would like to express my deepest gratitude to my advisor João Rosas, for his guidance, availability and support through this work. 
I would also like to acknowledge the Department of Electrical and Computer Engineering, the School of Science and Technology and the NOVA university for providing the academic environment and conditions that made not only this thesis, but my whole academic chapter possible.

I am also thankful to my family for their support, patience, and encouragement over the years. Their presence was essential, and this achievement would not have been possible without them.

Special thanks to all my friends, the ones with which I shared my academic adventures through five amazing years, and the ones who supported me through this chapter.

Finally, words cannot express my gratitude to to my girlfriend, for her constant support and motivation, as they were the main reason why this chapter is finally complete. 

Completing this thesis marks the end of a very important chapter in my life, and to anyone who was part of this path, my sincere thank you.

\end{ntacknowledgements}