%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% chapter1.tex
%% NOVA thesis document file
%%
%% Chapter with introduction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE chapter1.tex}%

\chapter{Introduction}
\label{cha:introduction}

\prependtographicspath{{Chapters/Figures/Covers/}}
\section{Context and Motivation}
In the last years, the digitalization of society has led to an increased
demand on \gls{IoT} systems. Whether the goal is equipment monitoring, traffic
management, security, or industrial automation, \gls{IoT} systems are a more
scalable and efficient alternative than the traditional methods.

\gls{IoT} systems are networks of \gls{IoT} devices connected to each other
that can collect, process, and transmit data in real-time.

Real-time monitoring requires processing efficiently a continuous flow of data from many
devices, resulting in a large volume of data. This data must be processed
quickly and accurately so the system can detect anomalies like equipment failures
efficiently.

IoT platforms face unique challenges like handling large volumes of
unstructured data, managing connectivity problems, and ensuring low-latency
processing. As the system grows, the number of \gls{IoT} devices connected to
the platform increases, introducing challenges related to scalability, data
security, and reliability.

To address these challenges, cloud architectures, allied to edge computing, provide highly scalable solutions
that allow efficient data collection, seamless device integration, and powerful
computational resources.

\gls{IoT} systems have also been relying more and more on \gls{AI} and \gls{ML} for
anomaly detection and predictive analytics, yet deploying complex \gls{AI}
models to resource-constrained \gls{IoT} devices remains a complex process.

\section{Problem}

As digitalization advances, monitoring and surveillance requirements are
becoming more and more complex, requiring scalable, highly customizable and
low-cost solutions.

Traditional monitoring platforms are often rigid and limited in their ability
to scale and adapt to new requirements, like new devices with different data.
The high processing capacity required for managing many \gls{IoT} devices,
ensuring real-time processing, and integrating predictive \gls{AI}/\gls{ML}
analytics emphasizes the need for more advanced architectures.

The increased digitalization in sectors like manufacturing, smart cities,
and agriculture requires an open, generic, and scalable platform that takes
advantage of the cloud and edge computing scaling benefits, is able to
adapt to diverse anomaly detection scenarios, and supports integration with
\gls{AI}/\gls{ML}.
\section{Proposed Solution}

The proposed solution is an open cloud-based \gls{IoT} platform for monitoring
and surveillance, complemented by an edge node for sites with unstable cloud
connectivity. It targets scenarios where heterogeneous devices send telemetry
continuously, abnormal conditions must be detected with low latency, and
operators need access to both live and historical system data. Typical
examples include equipment monitoring, building alarms, environmental sensing,
and traffic supervision.

The platform supports five user goals: monitoring devices in real time,
automating anomaly detection, reviewing and acknowledging alerts,
investigating device history, and continuing monitoring during temporary loss
of cloud connectivity. To support these goals, it provides user
authentication, device registration and provisioning, and telemetry ingestion
over \gls{MQTT}, \gls{HTTP}, and \gls{CoAP}. It also accepts binary file
uploads, discovers new telemetry fields automatically, supports rule-based and
statistical anomaly detection, and provides live dashboards, historical
queries, and configurable notifications.

The solution combines cloud services with local edge processing. In the cloud,
the platform offers centralized ingestion, storage, analytics, and web access
through an event-driven microservice architecture. At the edge, a local node
can ingest telemetry, evaluate the monitoring rules required for local
operation, store telemetry and alerts, and synchronize buffered data with the
cloud after connectivity is restored. This design addresses sites where
permanent cloud access cannot be assumed.

The main non-functional goals of the platform are low-latency monitoring,
support for heterogeneous payload structures, extensibility, offline
resilience, independent scaling of backend components, secure access for users
and devices, and preservation of telemetry and alert data during temporary
failures. These goals motivate the use of event-driven communication, which
allows new processing services, such as \gls{ETL}, business intelligence, or
\gls{AI}/\gls{ML} modules, to be added without changing the main ingestion
flow.

\section{Dissertation Outline}

This dissertation is organized into six chapters.

\begin{description}
	\item[Chapter~\ref{cha:State-Of-The-Art} -- State of the Art] reviews \gls{IoT} platforms, covering communication protocols, architectural patterns, edge computing, data storage, and related platforms. It identifies the gaps and limitations that motivate the design decisions in the following chapters.
	\item[Chapter~\ref{cha:ArchitectureAndDesign} -- Design and Architecture] defines the use cases, functional and non-functional requirements, and the overall system architecture. It describes the five-layer design, the event-driven communication model, the read-write path separation, the edge node architecture, and the data model.
	\item[Chapter~\ref{cha:Implementation} -- Implementation] details how the platform was built. It covers the technology stack, the backend service organization, deployment with Docker Compose, the security model, the database schema definitions, the implementation of each backend service and the edge node, the web application, and the observability stack.
	\item[Chapter~\ref{cha:Evaluation} -- Experimental Results and Evaluation] evaluates the platform through five experiments: a baseline performance study across three ingestion protocols, a horizontal scaling study on the \gls{MQTT} path, an offline resilience test, an extensibility validation with a Python vision service, and a device heterogeneity test with a real Raspberry Pi sensor.
	\item[Chapter~\ref{cha:Conclusions} -- Conclusions and Future Work] summarizes the main results, discusses the challenges encountered, and identifies directions for future work.
\end{description}
